\bf Range Analysis for the IonMonkey JavaScript engine

### 2012-04-29 project-writings/paper.tex
Write some intro stuff.

\section{Introduction}
%% XXX: FIXME, TODO
We propose implementing and evaluating range analysis for IonMonkey,
an optimizing JIT compiler for Javascript. Range analysis is
particularly useful in this context: in addition to taking advantage
of the analysis results to eliminate redundant comparisons and
bounds-checks, Javascript's dynamic semantics requires that all
numeric computations are floating point arithmetic.  This is slow, so
numbers that can be stored as integers are, and arithmetic operations
are guarded against overflow into their floating point
representation. With range analysis, we can prove these guards
unnecessary and emit higher-quality code.

>>>>

\section{Introduction}
With the explosion of browser based applications that has occured in
recent years, Javascript has become one of the most widely used
programming languages, and Javascript performance has become
critically important. Unfortunately, Javascript seems to be designed
to be un-performant.

One source of these woes are Javascript's dynamic semantics for
numbers: all numeric values are defined by the specification to be
IEEE 754 double precision floating point values (with bitwise
operations being defined to round to the nearest 32 bit integer before
performing the operation). Since the common case is that numeric
values are integers and because floating point operations are
more expensive, all major Javascript implementations attempt to
represent numbers as integers when possible.

In IonMonkey, Mozilla's next-generation optimizing Javascript JIT
compiler, an approximate type inference system is used to generate
type-specialized code for the common case of integer arithmetic, and
type guards are inserted if the type inferencer is not certain. The
catch is that since double precision floats have a larger range than
32-bit integers, if the integer operations would overflow the result
must be promoted to being stored in a double. IonMonkey deals with
this by inserting overflow checks after arithmetic operations and
bailing out of the type specialized code if they trip. While this is
sometimes actually necessary, it is frequently the case that overflow
is not possible.

We have implemented integer value range analysis in the IonMonkey
engine and used it to eliminate provably unnecessary overflow checks.
